---
title: "Getting Started with scRNA-R-Python-bridging"
author: "CC1001-CTRL"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: united
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                      fig.width = 8, fig.height = 6)
```

# Introduction

This vignette demonstrates how to use the scRNA-R-Python-bridging template for single-cell RNA sequencing analysis. The template includes essential packages like Seurat, SpatialExperiment, and ggplot2, along with a reproducible environment managed by renv.

## Package Installation

If you haven't already set up the environment, install the required packages:

```{r eval=FALSE}
# Install renv if not already installed
if (!requireNamespace("renv", quietly = TRUE)) {
  install.packages("renv")
}

# Restore the package environment
renv::restore()
```

## Load Required Libraries

```{r load_libraries}
library(Seurat)
library(ggplot2)
library(dplyr)
library(patchwork)

# Source our custom functions
source("../R/analysis_functions.R")
```

# Basic Single-Cell Analysis Workflow

## 1. Data Loading and Preprocessing

For this example, we'll use the built-in pbmc_small dataset that comes with Seurat:

```{r load_data}
# Load example data
data("pbmc_small")
pbmc <- pbmc_small

# Display basic information
pbmc
```

## 2. Quality Control

Let's examine the quality control metrics:

```{r qc_metrics}
# Add mitochondrial gene percentage
pbmc[["percent.mt"]] <- PercentageFeatureSet(pbmc, pattern = "^MT-")

# Create QC plots
qc_plots <- create_qc_plots(pbmc)
qc_plots$violin
```

```{r qc_scatter}
# Feature scatter plots
qc_plots$feature_scatter1
qc_plots$feature_scatter2
```

## 3. Normalization and Feature Selection

```{r normalization}
# Normalize and scale data
pbmc <- normalize_and_scale(pbmc, nfeatures = 2000)

# Show top variable features
top10 <- head(VariableFeatures(pbmc), 10)
print(top10)

# Plot variable features
var_plot <- VariableFeaturePlot(pbmc)
var_plot <- LabelPoints(plot = var_plot, points = top10, repel = TRUE)
var_plot
```

## 4. Dimensional Reduction

```{r dimensionality_reduction}
# Run PCA, clustering, and UMAP
pbmc <- run_dimensional_reduction(pbmc, npcs = 50)

# PCA plot
DimPlot(pbmc, reduction = "pca")
```

```{r umap}
# UMAP plot
DimPlot(pbmc, reduction = "umap", label = TRUE, pt.size = 0.5) + 
  ggtitle("UMAP of PBMC Clusters")
```

## 5. Cluster Analysis

```{r cluster_analysis}
# Number of cells per cluster
table(Idents(pbmc))

# Find marker genes for each cluster
markers <- FindAllMarkers(pbmc, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

# Display top markers for each cluster
top_markers <- markers %>%
  group_by(cluster) %>%
  slice_max(n = 3, order_by = avg_log2FC)

print(top_markers)
```

## 6. Visualization of Marker Genes

```{r marker_visualization}
# Plot expression of top marker genes
if(nrow(top_markers) > 0) {
  feature_genes <- head(unique(top_markers$gene), 6)
  FeaturePlot(pbmc, features = feature_genes, ncol = 3)
}
```

## 7. Spatial Analysis with SpatialExperiment

For spatial transcriptomics data, you can use SpatialExperiment:

```{r spatial_example, eval=FALSE}
library(SpatialExperiment)

# Example of creating a SpatialExperiment object
# (This is just a template - replace with your actual spatial data)
# spe <- SpatialExperiment(
#   assays = list(counts = your_count_matrix),
#   colData = your_cell_metadata,
#   spatialCoords = your_spatial_coordinates
# )
```

# Session Information

```{r session_info}
sessionInfo()
```

# Summary

This vignette demonstrated:

1. Loading and quality control of single-cell data
2. Normalization and feature selection
3. Dimensional reduction with PCA and UMAP
4. Clustering and marker gene identification
5. Visualization techniques

The template provides a solid foundation for single-cell RNA-seq analysis with reproducible package management through renv.

For more advanced analyses, refer to:
- [Seurat documentation](https://satijalab.org/seurat/)
- [SpatialExperiment vignettes](https://bioconductor.org/packages/SpatialExperiment/)
- [Bioconductor workflows](https://bioconductor.org/help/workflows/)